---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bgvars: Bayesian Inference for Global Vector Autoregressive (GVAR) and Global Vector Error Correction (GVEC) Models

[![CRAN status](https://www.r-pkg.org/badges/version/bgvars)](https://cran.r-project.org/package=bgvars)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/bgvars)](https://cran.r-project.org/package=bgvars)
[![Total downloads](https://cranlogs.r-pkg.org/badges/grand-total/bgvars)](https://cran.r-project.org/package=bgvars)
[![R-CMD-check](https://github.com/franzmohr/bgvars/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/franzmohr/bgvars/actions/workflows/R-CMD-check.yaml)
[![License: GPL (>= 3)](https://img.shields.io/badge/license-GPL%20%28%3E%3D%203%29-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)

[![GitHub Sponsors](https://img.shields.io/badge/Sponsor-%E2%9D%A4-ea4aaa?logo=github-sponsors&logoColor=white)](https://github.com/sponsors/franzmohr)
[![Buy Me a Coffee](https://img.shields.io/badge/Buy%20Me%20a%20Coffee-FFDD00?logo=buymeacoffee&logoColor=black)](https://www.buymeacoffee.com/franzmohr)


## Overview

The `bgvars` package provides functions for the specification, estimation and evaluation of Bayesian global autoregressive (GVAR) models. Global vector autoregressive (GVAR) models are convenient tools to model the world economy. They were originally proposed by Pesaran et al. (2004) and further developed by Dees et al. (2007) to study the international transmission of shocks. Since then they have been applied to a range of macroeconomic topics. Chudik and Pesaran (2016) provide an extensive survey of the latest developments in GVAR modelling.

## Installation

```{r cran, include = FALSE, eval = FALSE}
install.packages("bgvars")
```

### Development version

```{r github, eval = FALSE}
# install.packages("devtools")
devtools::install_github("franzmohr/bgvars")
```

## Usage

The `bgvars` package allows Bayesian inference of GVAR models. It separates a typical GVAR analysis into four steps:

* *Preparation*, which includes the initial transformation of raw data and the generation of regional aggregates as well as the weight matrices $W_{i, t}$;
* *Specification* for setting up country-specific models that should be estimated;
* *Estimation* using Bayesian algorithms to produce draws from the posterior distribution of each country model, which are subsequently combined to a global model;
* *Evaluation* for the generation of forecasts, impulse responses and forecast error variance decompositions.

### Data

The `bgvars` packages comes with the updated GVAR database of Mohaddes and Raissi (2020), which contains economic time series for 33 countries and 3 commodities from 1979Q2 to 2019Q4.[^gvar2019]

```{r data, eval = TRUE}
#library(bgvars)
devtools::load_all(".")
data("gvar2023")

submodel_data <- gvar2023[["submodel_data"]] # Country series including weights
global_data <- gvar2023[["global_data"]] # Global commodities data
region_weights <- gvar2023[["region_weights"]] # Data for regional weights



# Take first differences of non-stationary series  
#submodel_data <- diff(submodel_data, variables = c("p"), multi = 100)
```

### Preparation

Regional series can be calculated from individual submodel series with the function `create_regions`. It calculates a region based on the time series specified in `region_weights`. The resulting object is of class `submodeldata` and contains an updated version of `submodel_data`.

```{r region, eval = TRUE}
# Generate EA area region with 3 year rolling window weights
ea <- c("AT", "BE", "DE", "ES", "FI", "FR", "IT", "NL")
submodel_data <- create_regions(submodel_data = submodel_data,
                                regions = list("EA" = ea),
                                period = 3,
                                region_weights = region_weights)
```

```{r initialise, eval = TRUE}
object <- create_gvarmodel(submodel_data = submodel_data,
                           global_data = global_data)
```

Weight matrices for each country/region can be calculated with the function `add_weight_matrices`.


```{r weights, eval = TRUE}
# Generate weight matrices as 3 year, rolling window averages
object <- add_weight_matrices(object = object,
                              submodel_data = submodel_data,
                              period = 3)
```


### Model specification

```{r, eval = TRUE}
# Create an object with country model specifications
object <- add_submodels(object,
                        endogen = c("y", "Dp"),
                        p_endogen = 0:2,
                        exogen = c("y", "Dp"),
                        p_exogen = 0:1,
                        global = "poil",
                        s = 0,
                        deterministic = "const",
                        error = "wishart",
                        iterations = 3000, burnin = 1000)
```

<!-- `bgvars` assists in the set-up country models by producing a list, where each element contains all the information required to estimate a model. First, the function `create_specifications` produces an object, which contains the specifications for each country. -->


<!-- Country specifications can be changed by accessing the elements of the list directly: -->

<!-- ```{r, eval = TRUE} -->
<!-- model_specs$US$domestic$variables <- c("y", "Dp", "r") -->
<!-- model_specs$US$foreign$variables <- c("y", "Dp") -->
<!-- ``` -->

<!-- `create_models` produces all country models, which should be estimated. This allows for easy parallelisation. The number of elements of the resulting list depends on the specification of domestic lags, foreign lags, global lags, and the rank of the cointegration matrix `r`, if sepecified. For example, if the lags of domestic variables were set to `lags = 1:2`, the function produces two country submodels for each specification. -->

<!-- ```{r, eval = TRUE} -->
<!-- # Create estimable objects -->
<!-- object <- create_models(country_data = country_data, -->
<!--                         weight_data = gvar_weights, -->
<!--                         global_data = global_data, -->
<!--                         model_specs = model_specs) -->
<!-- ``` -->

Finally, priors must be specified with `add_priors`.

```{r, eval = TRUE}
object <- add_priors(object,
                     coef = list(v_i = 1),
                     sigma = list(df = 3, scale = 0.0001))
```

```{r, eval = TRUE}
object <- add_initial_values(object)
```

### Estimate model

```{r, eval = TRUE, message=FALSE, results='hide'}
object <- add_posterior_coefficients(object)
```

### Model selection


```{r, eval = TRUE, message=FALSE, results='hide'}
object <- add_posterior_loglik(object)
```


```{r, eval = TRUE, message=FALSE, results='hide'}
sc <- add_posterior_loglik(object)
```


The estimated country models can be combined and solved with the function `combine_submodels`.

```{r, message=FALSE, results='hide'}
gvar <- combine_submodels(object)
```

### Impulse response analysis

Impulse response analysis can be done with the `girf` function.

```{r, eval = TRUE, fig.width=5.5, fig.height=4.5}
gvar_irf <- girf(gvar, impulse = c("US", "r"),
                 response = c("EA", "y"),
                 n.ahead = 20, ci = .68)

plot(gvar_irf)
```


## References

Chudik, A. & Pesaran, M. H. (2016). Theory and practice of GVAR modelling. *Journal of Economic Surveys 30*(1), 165-197. <https://doi.org/10.1111/joes.12095>

Dees, S., Mauro, F., Pesaran, M. H., & Smith, V. L. (2007). Exploring the international linkages of the euro area: A global VAR analysis. *Journal of Applied Econometrics 22*(1), 1-38. <https://doi.org/10.1002/jae.932>

George, E. I., Sun, D., & Ni, S. (2008). Bayesian stochastic search for VAR model restrictions. *Journal of Econometrics, 142*(1), 553-580. <https://doi.org/10.1016/j.jeconom.2007.08.017>

Koop, G., Pesaran, M. H., & Potter, S.M. (1996). Impulse response analysis in nonlinear multivariate models. *Journal of Econometrics 74*(1), 119-147. <https://doi.org/10.1016/0304-4076(95)01753-4>

Lütkepohl, H. (2007). *New introduction to multiple time series analysis* (2nd ed.). Berlin: Springer.

Mohaddes, K., & Raissi, M. (2020). Compilation, revision and updating of the global VAR (GVAR) database, 1979Q2--2019Q4 (mimeo). <https://www.mohaddes.org/gvar>.

Pesaran, H. H., & Shin, Y. (1998). Generalized impulse response analys is in linear multivariate models. *Economics Letters, 58*(1), 17-29. <https://doi.org/10.1016/S0165-1765(97)00214-0>

Pesaran, M., Schuermann, T., & Weiner, S. M. (2004). Modeling regional interdependencies using a global error-correcting macroeconometric model. *Journal of Business & Economic Statistics 22*(2), 129-162. <https://doi.org/10.1198/073500104000000019>

[^gvar2019]: The paper and data set can be downloaded from <https://www.mohaddes.org/gvar>.
