---
title: "Introduction to `bgvars`"
author: "Franz X. Mohr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to bgvars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Global vector autoregressive (GVAR) models are convenient tools to model the world economy. They were originally proposed by Pesaran et al. (2004) and further developed by Dees et al. (2007) to study the international transmission of economic shocks. Since then they have been used in a range of macroeconomic applications. Chudik and Pesaran (2016) provide an extensive survey of the latest developments in GVAR modelling. The `bgvars` package allows Bayesian inference of such models.

GVAR models address the so-called *curse of dimensionality* of standard VAR models, where the number of estimated coefficients increases disproportionately with the number of variables. Since macroeconomic time series tend to be relatively short, there might not be enough observations for the estimation of large VAR models. This problem is especially severe in the realm of international spill-over effects, where the interaction of variables across many countries are of interest. Based on a few assumptions the GVAR approach tries to mitigate this challenge by dividing such an estimation problem into multiple, country-specific models. For a basic GVAR model with global variables and constant coefficients such a model can have the following form

$$y_{i,t} = \sum_{l = 1}^{p^d_i} A^{d}_{i,l} y_{i, t - l} +
\sum_{l = 0}^{p_i^*} A^{*}_{i,l} y^*_{i, t - l}+
\sum_{l = 0}^{s_i} B_{i,l} x_{t - l} +
C_i d_{i,t} + u_{i,t},$$

where $y_{i,t}$ is a $K_i$-dimensional vector of endogenous domestic variables of country $i$ in period $t$ and $K_i$ is the number of endogenous domestic variables for that country. $y^{*}_{i,t}$ is a $K^{*}_{i} \times 1$ vector of country-specific foreign variables. $A^{d}_{i,l}$ is a $K_i \times K_i$ coefficient matrix of lagged values of endogenous domestic variables. $A^{*}_{i,l}$ is the $K_i \times K^{*}_{i}$ coefficient matrix of foreign variables. $p^{d}_{i}$ and $p^{*}_{i}$ are the lag orders of endogenous and foreign variables, respectively. $B_{i,l}$ is a $K_i \times K^{g}_{i}$ coefficient matrix corresponding to the $K^{g}_{i} \times 1$ vector $x_t$ of global variables. $C_{i}$ is the coefficient matrix of deterministic terms in vector $d_{i,t}$. The error term $u_t$ is assumed to have zero mean and a variance-covariance matrix $\Sigma_{i}$.

Subtracting the term of contemporaneous foreign variables and using $v_{i, t} = (y^{\prime}_{i,t}, y^{*\prime}_{i,t})^{\prime}$ gives

$$\left[I_{K_i}, -A_{i,0}^{*} \right]  v_{i,t} =
\sum_{l = 1}^{p_i}  \left[ A^d_{i,l}, A_{i,l}^* \right] v_{i,t-l} +
\sum_{l = 0}^{s_i} B_{i, l} x_{t - j} +
C_i d_{i,t} + u_{i, t},$$

where $A^{d}_{i,l}$ and $A^{*}_{i,l}$ are set to zero if $p^{d}_{i}$ or $p^{*}_{i}$ are smaller than $p_i$, respectively.

The country-specific foreign variables in $y^{*}_{i,t}$ are constructed as weighted averages of foreign observations. Therefore, the vector of country-specific domestic and foreign variables can be written in terms of the global $K$-dimensional vector $y_t$ so that $v_{i, t} = W_{i} y_t$, where $W_{i}$ is a $K_i \times K$ weight matrix and $K = \sum_{i = 1}^{N} K_i$ is the total number of endogenous variables in the global model. Defining $A_{i,0} = \left[I_{K_i}, -A^{*}_{i,0}\right] W_{i, t}$ and $A_{i,l} = \left[A^{d}_{i,l} A^{*}_{i,l}\right] W_{i, t}$ for $l = 1, ..., p_i$ each country model can be written as

$$A_{i,0} y_t =
\sum_{l = 1}^{p_{i}} A_{i, l} y_{t-l} +
\sum_{l = 0}^{s_{i}} B_{i, l} x_{t-l} +
C_{i} d_{i,t} +
u_{i,t}.$$

Note that the country-specific weight matrix is one of the central aspect of GVAR modelling. The matrix helps to reduce the dimensionality of the overall model by adding prior information about the relative importance of a foreign country to the model.

Stacking the country-specific coefficient matrices gives the *structural GVAR model*

$$A_{0} y_t = \sum_{l = 1}^{p} A_{l} y_{t-l} + \sum_{l = 1}^{s} B_{l} x_{t} + C d_t + u_t,$$
where $p = max(p^d_1,...,p^d_N, p^{*}_1,...,p^{*}_N)$ is the lag of the global system, $s = max(s_1,...,s_N)$ is the lag of the global variable and $u_t$ is normally distributed with zero mean and covariance matrix $\Sigma$. Furthermore,

$$A_j = \begin{bmatrix} A_{1, j} \\ A_{2,j} \\ \vdots \\ A_{N,j} \end{bmatrix},
B_j = \begin{bmatrix} B_{1, j} \\ B_{2,j} \\ \vdots \\ B_{N,j} \end{bmatrix},
C = \begin{bmatrix} C_{1} \\ C_{2} \\ \vdots \\ C_{N} \end{bmatrix},
\text{ and } \Sigma = \begin{bmatrix} \Sigma_1 & 0 & \dots & 0 \\ 0 & \Sigma_2 & \ddots & \vdots \\ \vdots & \ddots & \ddots & 0 \\ 0 & \dots & 0 & \Sigma_{N} \end{bmatrix}.$$

The *reduced form GVAR model* is

$$y_t = \sum_{l = 1}^{p} G_{l} y_{t-l} + \sum_{l = 1}^{s} H_{l} x_{t-l} + D d_t + A_0^{-1} u_t,$$
where $G_l = A_0^{-1} A_l$, $H_l = A_0^{-1} B_l$ and $D = A_0^{-1} C$. This model resembles a standard VAR model with exogenous variables, for which impulse responses and variance decompositions can be calculated in the usual manner.

## Data

The `bgvars` package comes with the *2016 Vintage* of the GVAR dataset (Mohaddes and Raissi, 2018), which contains economic time series for 33 countries and three commodities from 1979Q2 to 2016Q4.

```{r data, eval = FALSE}
library(bgvars) 
data("gvar2016")
```

### Country and commodity data

```{r country data, eval = FALSE}
country_data <- gvar2016$country_data
global_data <- gvar2016$global_data
```


### Regional data

In Dees et al. (2007) the observations of eight euro area countries are combined to obtain a time series for the euro area regiona. The function `create_regions` can be used to do this and the data set `region_weights` in `gvar2016` contains suitable weight data.

```{r regions, eval = FALSE}
region_weights <- gvar2016$region_weights

# Define region for the Euro area
ea <- c("AT", "BE", "DE", "ES", "FI", "FR", "IT", "NL")

temp <- create_regions(country_data = country_data,
                       regions = list("EA" = ea),
                       period = 3,
                       region_weights = region_weights,
                       gvar_weights = gvar_weights)

# Update country and weight data
country_data <- temp$country_data
gvar_weights <- temp$gvar_weights
```

### Weight matrix data



```{r weights, eval = FALSE}
weight_data <- gvar2016$weight_data

gvar_weights <- create_weights(weight_data = weight_data, period = 3,
                               country_data = country_data)
```


## Model specification

```{r specs, eval = FALSE}
model_specs <- create_specifications(country_data = country_data,
                                     global_data = global_data,
                                     p_domestic = 2,
                                     p_foreign = 1,
                                     r = 2,
                                     countries = c("EA", "US", "JP", "CA", "MX", "GB"),
                                     variables = c("y", "Dp", "r", "poil"),
                                     type = "VEC")
```


```{r us, eval = FALSE}
model_specs$US$domestic$variables <- c("poil", "y", "Dp", "r")
model_specs$US$foreign$variables <- c("y", "Dp")
model_specs$US$cointegration$rank <- 1

model_specs$MX$domestic$lags <- 1
```

The function `create_models` is used to produce a list of models that can be estimated. This list structure was chosen to facilitate the use of parallel computing, where each element of that list - a country model - can be estimated on a seperate node.

```{r models, eval = FALSE}
country_models <- create_models(country_data = country_data,
                                gvar_weights = gvar_weights,
                                global_data = global_data,
                                model_specs = model_specs)
```

Note that once a global variable enters a country model as endogenous variable, `create_models` automatically turns it into a foreign star variable in other countries and excludes it from the set of global variables.

Deterministic terms can be added to each country model with the `add_deterministics` function. For VEC models it can be chosen, whether constants, linear trends or seasonal dummy variables enter the error correction term as restricted variables or the non-cointegration part of the model as unrestricted variables. Similar to the existing literature, a trend is added to the error correction term and a constant to the non-cointegration part of the model.

```{r deterministics, eval = FALSE}
country_models <- add_deterministics(country_models,
                                     const = "unrestricted",
                                     trend = "restricted")
```

## Priors

Before the country models can be estimated, the priors of the coefficients have to be specified. The `add_priors` function can 

```{r priors, eval = FALSE}
country_models <- add_priors(country_models)
```

## Estimation

The country-specific models can be estimated with the `estimate_gvar` function.

```{r estimate show, message = FALSE, include = FALSE, eval = FALSE}
gvar_est <- estimate_gvar(country_models, iterations = 25000, burnin = 10000, thin = 5)
```
```{r estimate, message = FALSE, eval = FALSE}
gvar_est <- estimate_gvar(country_models, iterations = 25000, burnin = 10000, thin = 5)
```


```{r contemp, eval = FALSE}
contemp_coeffs(gvar_est)
```

## Solving the GVAR model

Once the country-specific models are estimated.

```{r, message=FALSE, include = FALSE, eval = FALSE}
gvar <- solve_gvar(gvar_est)
```
```{r, message=FALSE, eval = FALSE}
gvar <- solve_gvar(gvar_est)
```

## Impulse response analysis

```{r gir, eval = FALSE, fig.width=5.5, fig.height = 4.5}
gvar_irf <- girf(gvar, impulse = c("US", "r"), response = c("US", "y"),
                n.ahead = 20, ci = .68)

plot(gvar_irf * 100)
```

## Forecast error variance decomposition

```{r gfevd, eval = FALSE, fig.width=5.5, fig.height = 4.5}
gvar_fevd <- gfevd(gvar, response = c("US", "y"), n.ahead = 20, normalise_gir = TRUE)

plot(gvar_fevd)
```

## References

Chudik, A. & Pesaran, M. H. (2016). Theory and practice of GVAR modelling. *Journal of Economic Surveys 30*(1), 165-197. <https://doi.org/10.1111/joes.12095>

Dees, S., Mauro, F., Pesaran, M. H., & Smith, V. L. (2007). Exploring the international linkages of the euro area: A global VAR analysis. *Journal of Applied Econometrics 22*(1), 1-38. <https://doi.org/10.1002/jae.932>

Koop, G., Pesaran, M. H., & Potter, S.M. (1996). Impulse response analysis in nonlinear multivariate models. *Journal of Econometrics 74*(1), 119-147. <https://doi.org/10.1016/0304-4076(95)01753-4>

Lütkepohl, H. (2007). *New introduction to multiple time series analysis* (2nd ed.). Berlin: Springer.

Mohaddes, K., & Raissi, M. (2018). Compilation, revision and updating of the global VAR (GVAR) database, 1979Q2--2016Q4 (mimeo). <https://www.mohaddes.org/gvar>.

Pesaran, H. H., & Shin, Y. (1998). Generalized impulse response analys is in linear multivariate models. *Economics Letters, 58*(1), 17-29. <https://doi.org/10.1016/S0165-1765(97)00214-0>

Pesaran, M., Schuermann, T., & Weiner, S. M. (2004). Modeling regional interdependencies using a global error-correcting macroeconometric model. *Journal of Business & Economic Statistics 22*(2), 129-162. <https://doi.org/10.1198/073500104000000019>