---
title: "Global Vector Error Correction Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Global Vector Error Correction Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Data

The `bgvars` packages comes with the updated GVAR database of Mohaddes and Raissi (2020), which contains economic time series for 33 countries and 3 commodities from 1979Q2 to 2019Q4.[^gvar2019]

```{r data, message=FALSE, eval = TRUE}
#library(bgvars)
devtools::load_all(".")
data("dees2007")

submodel_data <- dees2007[["submodel_data"]]
global_data <- dees2007[["global_data"]]
```

```{r poil-into-us-model, message=FALSE, eval = TRUE}
# Add global series to US series
vars_us <- c(dimnames(submodel_data[["US"]][["endogen"]])[[2]], "poil")
submodel_data[["US"]][["endogen"]] <- cbind(submodel_data[["US"]][["endogen"]], global_data)
dimnames(submodel_data[["US"]][["endogen"]])[[2]] <- vars_us
```


### Preparation

```{r initialise, eval = TRUE}
object <- create_gvecmodel(submodel_data = submodel_data)
```

Weight matrices for each country/region can be calculated with the function `add_weight_matrices`.


```{r weights, eval = TRUE}
# Generate weight matrices
object <- add_weight_matrices(object = object,
                              submodel_data = submodel_data,
                              period = 1999:2001)
```


### Model specification

To reproduce the results of Dees et al. (2007), each sub-model is added manually for each country based on the specifications provided in the article's supplementary material.

```{r}
submodel <- "US"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "rl", "q", "poil"),
  p_endogen = 2,
  exogen = c("y", "Dp", "ep"),
  p_exogen = 2,
  r = 2,
  const = "unrestricted",
  trend = "restricted"
)
```


```{r}
submodel <- "EA"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "rl", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 2,
  r = 2,
  const = "unrestricted",
  trend = "restricted"
)
```


```{r}
submodel <- "CN"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  r = 1,
  const = "unrestricted",
  trend = "restricted"
)
```


```{r}
submodel <- "JP"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "rl", "q", "ep"),
  p_endogen = 1,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  r = 4,
  const = "unrestricted",
  trend = "restricted"
)
```


```{r}
submodel <- "GB"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "rl", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  r = 3,
  const = "unrestricted",
  trend = "restricted"
)
```


```{r}
submodel <- "SE"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "rl", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  r = 3,
  const = "unrestricted",
  trend = "restricted"
)
```


```{r}
submodel <- "CH"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "rl", "q", "ep"),
  p_endogen = 1,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  r = 3,
  const = "unrestricted",
  trend = "restricted"
)
```


```{r}
submodel <- "NO"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "rl", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  r = 2,
  const = "unrestricted",
  trend = "restricted"
)
```


```{r}
submodel <- "AR"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 2
)
```


```{r}
submodel <- "BR"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 1
)
```


```{r}
submodel <- "CL"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 2
)
```


```{r}
submodel <- "MX"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "ep"),
  p_endogen = 1,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 3
)
```


```{r}
submodel <- "PE"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 3
)
```


```{r}
submodel <- "AU"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "rl", "q", "ep"),
  p_endogen = 1,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 4
)
```


```{r}
submodel <- "CA"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "rl", "q", "ep"),
  p_endogen = 1,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 4
)
```


```{r}
submodel <- "NZ"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "rl", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 3
)
```


```{r}
submodel <- "ID"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 3
)
```


```{r}
submodel <- "KR"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "rl", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 4
)
```


```{r}
submodel <- "MY"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 1
)
```


```{r}
submodel <- "PH"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 2
)
```


```{r}
submodel <- "SG"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 3
)
```


```{r}
submodel <- "TH"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "q", "ep"),
  p_endogen = 1,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 3
)
```


```{r}
submodel <- "IN"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 2
)
```



```{r}
submodel <- "ZA"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "rl", "q", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 1
)
```


```{r}
submodel <- "SA"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 1
)
```


```{r}
submodel <- "TR"
object[["submodels"]][[submodel]] <- create_vecxsubmodel(object,
  submodel = submodel,
  endogen = c("y", "Dp", "rs", "ep"),
  p_endogen = 2,
  exogen = c("y", "Dp", "q", "rs", "rl", "poil"),
  p_exogen = 1,
  const = "unrestricted",
  trend = "restricted",
  r = 1
)
```


The resulting sub-models have different sample sizes. Function `align_model_obs` can be employed to ensure that all sub-models use the same set of observations that are common to all sub-models.

```{r}
object <- align_model_obs(object)
```



### Prior specification

```{r, eval = TRUE}
object <- add_priors(object,
                     coef = list(v_i = 0),
                     coint = list(v_i = 0, p_tau_i = 1),
                     sigma = list(df = 3, scale = 0.0001))
```

### Initial values

```{r, eval = TRUE}
object <- add_initial_values(object)
```

Since the function creates initial values based on the reduced rank maximum likelihood estimator, closer inspection of the sub-models should show that the produced estimates are equal to the values provided in the supplementary material to Dees et al. (2007).

### Posterior simulation

#### BayesTS

Given the weight matrices, the sub-models of a global model are independent of
one another, which makes it possible to estimate them outside R.
`write_to_hdf5` exports the model to a folder, one HDF5 file per sub-model
beside a `model.h5` holding the global data, the weight matrices and a table
describing every sub-model file.

```{r, eval = FALSE}
write_to_hdf5(object, folder = "C:/path/to/gvar")
```

[BayesTS](https://github.com/franzmohr/BayesTS) is then pointed at the folder of
sub-models and writes the posterior draws back into the files it read.

```{bash, eval = FALSE}
bayests posterior C:/path/to/gvar/submodels
```

`read_gvar_from_folder` brings the whole model back, draws included. The
sub-models are found through the table in `model.h5` rather than through the
directory tree, so the names they are known by survive a folder that has been
moved or renamed.

```{r, eval = FALSE}
object <- read_gvar_from_folder("C:/path/to/gvar")
```


#### Package 

```{r, eval = TRUE, message=FALSE, results='hide'}
object <- add_posterior_coefficients(object)
```

```{r, eval = TRUE, message=FALSE, results='hide'}
object <- add_posterior_loglik(object)
```

### Model selection


```{r, eval = TRUE, message=FALSE, results='hide'}
sc <- selection_criteria(object)
```


The estimated country models can be combined and solved with the function `combine_submodels`.

```{r, message=FALSE, results='hide'}
gvar <- combine_submodels(object)
```

### Impulse response analysis

Impulse response analysis can be done with the `girf` function.

```{r, eval = TRUE, fig.width=5.5, fig.height=4.5}
gvar_irf <- girf(gvar, impulse = c("US", "r"),
                 response = c("EA", "y"),
                 n.ahead = 20, ci = .68)

plot(gvar_irf)
```


## References

Chudik, A. & Pesaran, M. H. (2016). Theory and practice of GVAR modelling. *Journal of Economic Surveys 30*(1), 165-197. <https://doi.org/10.1111/joes.12095>

Dees, S., Mauro, F., Pesaran, M. H., & Smith, V. L. (2007). Exploring the international linkages of the euro area: A global VAR analysis. *Journal of Applied Econometrics 22*(1), 1-38. <https://doi.org/10.1002/jae.932>

George, E. I., Sun, D., & Ni, S. (2008). Bayesian stochastic search for VAR model restrictions. *Journal of Econometrics, 142*(1), 553-580. <https://doi.org/10.1016/j.jeconom.2007.08.017>

Koop, G., Pesaran, M. H., & Potter, S.M. (1996). Impulse response analysis in nonlinear multivariate models. *Journal of Econometrics 74*(1), 119-147. <https://doi.org/10.1016/0304-4076(95)01753-4>

Lütkepohl, H. (2007). *New introduction to multiple time series analysis* (2nd ed.). Berlin: Springer.

Mohaddes, K., & Raissi, M. (2020). Compilation, revision and updating of the global VAR (GVAR) database, 1979Q2--2019Q4 (mimeo). <https://www.mohaddes.org/gvar>.

Pesaran, H. H., & Shin, Y. (1998). Generalized impulse response analys is in linear multivariate models. *Economics Letters, 58*(1), 17-29. <https://doi.org/10.1016/S0165-1765(97)00214-0>

Pesaran, M., Schuermann, T., & Weiner, S. M. (2004). Modeling regional interdependencies using a global error-correcting macroeconometric model. *Journal of Business & Economic Statistics 22*(2), 129-162. <https://doi.org/10.1198/073500104000000019>

[^gvar2019]: The paper and data set can be downloaded from <https://www.mohaddes.org/gvar>.
